---
title: "Title of Blog Post Deliverable"
author: 'Xinyu Gao'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
    toc: yes
    number_sections: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction 

This assignment is going to explore how to use analytic and simulation ways to develop functions for the distribution of order statistics. For the generated density function, using the histogram to test it; for the generated probability function, using the ecdf to test it; for the generated quantile function, using QQ plot to test it. It also explores how to develop previously defined density, probability, and quantile functions into more advanced versions that can be applied to any distribution.

## Background

In this assignment, we need to know the formula for how to generate the density and probability function of order statistics, which will be mentioned in detail below. Also, we need to know how to generate quantile function from the probability function. I use the approfun() function to achieve it. For the testing part, we need to use histogram, ecdf, and QQ plot to achieve tasks. We need to use "..." to update density, probability, and quantile functions to make them able to deal with various kinds of distributions. 

# Methods
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(tgsify))
```

## Question 1
$$f_k(x)=k{n \choose k}[F(x)]^{k−1}[1−F(x)]^{n−k}f(x)$$
Based on the question introduction, n=200,k=100(median), and this is about the standard normal distribution
```{r}
# according to the above formula (k = 100, n = 200)
dorder <- function(x){
  100*
    choose(200,100)*
    (pnorm(x, mean=0, sd =1))^(100-1)*
    (1-pnorm(x, mean=0, sd =1))^(200-100)*
    dnorm(x, mean=0, sd =1)
}
# plot of the function
curve(
  dorder(x),
  -1,
  1,
  ylab = "Density",
  main = "PDF Of Standard Normal Distribution's Median"
)
```
\
This is the PDF of the standard normal distribution' median

## Question 2
$$F_{k}(x)=\sum_{j=k}^{N}{n \choose j}[F(x)]^j[1−F(x)]^{n−j}$$
Based on the question introduction, n=200,k=100, and this is about the standard normal distribution
```{r}
# according to the above formula(k = 100, n = 200)
porder <- function(x){
  pbinom(100-1, 200, pnorm(x, mean=0, sd=1), lower.tail = FALSE)
}
curve(
  porder(x),
  -1,
  1,
  ylab = "Probability",
  main = "CDF Of Standard Normal Distribution's Median"
)
```
\
This is the CDF of the standard normal distribution' median

## Question3
"qorder" is the inverse of "porder" 
\
Step1: create a look up table with value and its corresponding probability. The probability is calculated using "porder"
\
Step2: find the approximate function of the relationship between probability and value using "approxfun". In this function, probability is the input and value is the output
```{r}
qorder <- function(x){
  # look up table
  lt <- data.frame(x=seq(-4,4,by=0.01)) %>%
  mutate(p = porder(x)) 
  # function of the relationship between probability and value
  qf <-approxfun(lt$p,lt$x,ties="ordered")
  # using this function to find corresponding value given probability x
  qf(x)
}

curve(qorder(x),
      0,
      1,
      ylab= "x",
      xlab="quantile",
      ylim = c(-0.5,0.5),
      main = "Quantile Of Standard Normal Distribution's Median")
```
\
This is the quantile function of standard normal distribution' median, which is actually the inverse of the CDF

## Question 4
Use "ecdf" and "porder" function to finish this question
```{r}
# simulation
median_norm <- rep(NA,10000)
for (i in seq_along(median_norm))
{
  # calculate median for each sample
  median_norm[i] = rnorm(200,mean = 0, sd=1) %>% median
}
# the output is a function
bmi_ecdf <- ecdf(median_norm)  
plot(bmi_ecdf, do.points = FALSE, col = "red", lwd = 3, main="ecdf VS CDF")
# use porder to get the plot of the CDF
curve(porder(x),lwd = 3, col = "blue", add=TRUE)

# define the property of the legend
legend(
    "topleft"
  , c("ECDF","CDF")
  , lwd = 3
  , col = c("red","blue")
  , bty = "n"
)
```
\
Based on the graphic, the line of ecdf is basically consistent with the line of the generated probability function, which proves that the probability function is likely to be accurate.

## Question 5
Generate histogram of median sampling distribution and use "dorder" to get the theoretical PDF
```{r}
# get 10000 sample
median_norm <- rep(NA,10000)
for (i in seq_along(median_norm))
{
  # calculate median for each sample
  median_norm[i] = rnorm(200,mean = 0, sd=1) %>% median
}
# plot 10000 sample
hist(
  median_norm,
  breaks = 100,
  main = "Sampling distribution VS PDF",
  xlab = "x",
  freq = FALSE
)
# use "dorder" to get the plot of PDF
curve(
  dorder(x),
  lwd = 3, 
  col = "blue",
  add=TRUE
)
```
\
Based on the graphic, the shape of histogram is basically consistent with the shape of the generated density function，which proves that the density function is likely to be accurate.

## Question 6
```{r}
# get 10000 sample
median_norm_2 <- rep(NA,10000)
for (i in seq_along(median_norm_2)){
  # calculate median for each sample
  median_norm_2[i]  <- rnorm(200,mean=0, sd=1) %>% median
}
# sort the list
median_norm_2 <- median_norm_2 %>% sort()
# want to explore 199 quantiles except 0 and 1 quantile
y <- rep(NA, 199)
for (i in seq_along(y))
{
  y[i] = median_norm_2[10000*i/200]
}
# compatible with x
x = qorder((1:199)/200)

plot(x,y
     ,asp=1
     ,main = "Median Theoretical Quantile VS Sample Quantile"
     ,xlab = "Theoretical quantile"
     ,ylab = "Sample quantile"
     )
abline(0,1)

```
\
Based on the graphic, the plotted points almost fall along the line y=x. Thus the simulated median agree with the theoretical sampling distribution, which proves that the quantile function is likely to be accurate.

## Question 7
The updated function is the version of corresponding function from Question 1 using k to replace 100.
```{r}
# n=200,k=k
dorder_anyOrder <- function(x,k){
  k*
    choose(200,k)*
    (pnorm(x, mean=0, sd =1))^(k-1)*
    (1-pnorm(x, mean=0, sd =1))^(200-k)*
    dnorm(x, mean=0, sd =1)
}

# n=200,k=k
porder_anyOrder <- function(x,k){
  pbinom(k-1, 200, pnorm(x, mean=0, sd=1), lower.tail = FALSE)
}

# n = 200,k=k
qorder_anyOrder <- function(x,k){
  lt <- data.frame(x=seq(-4,15,by=0.01)) %>%
     mutate(p = porder_anyOrder(x,k)) 
  qf <-approxfun(lt$p,lt$x, ties ="ordered")
  qf(x)
}
```

## Question 8
```{r}
# get 10000 samples
max_norm <- rep(NA,10000)
for (i in seq_along(max_norm)){
  # calculate the max for each sample
  max_norm[i] = rnorm(200,mean=0, sd=1) %>% max
}
# sort the list
max_norm <- max_norm %>% sort()

# want to explore 199 quantiles
y <- rep(NA,199)
for (i in seq_along(y)){
  y[i] = max_norm[10000*i/200]
}
# use function "qorder_anyOrder" with k = 200(max)
x<-qorder_anyOrder((1:199)/200,200)
plot(x,y, asp = 1,
     xlab = "Theoretical quantile", 
     ylab = "Sample quantile",
     main =  "Max Theoretical Quantile VS Sample Quantile")
abline(0,1)
```
\
Based on the above graphic, the plotted points almost fall along the line y=x. Thus the simulated max agree with the theoretical sampling distribution, which proves that the updated quantile function is likely to be accurate.

## Question 9
The new function needs two more variables "dist" and  "...". "dist" contains the name information of the distribution and "..." contains the parameter information of the distribution
```{r}
# combine two strings
`%|%` <- function(a,b) paste0(a,b)

dorder_anyOrderDistri <- function(x,k,dist,...){
  # CDF
  F <- eval(parse(text ="p" %|% dist))
  # PDF
  f <- eval(parse(text ="d" %|% dist))
  k*
    choose(200,k)*
    (F(x, ...))^(k-1)*
    (1-F(x, ...))^(200-k)*
    f(x, ...)
}

porder_anyOrderDistri <- function(x,k,dist,...){
  # CDF
  F <- eval(parse(text ="p" %|% dist))
  pbinom(k-1, 200, F(x, ...), lower.tail = FALSE)
}

qorder_anyOrderDistri <- function(x,k,dist,...){
  lt <- data.frame(x=seq(-4,15,by=0.01)) %>%
     mutate(p = porder_anyOrderDistri(x,k,dist,...)) 
  qf <-approxfun(lt$p,lt$x)
  qf(x)
}
```

## Question 10
Based on the question introduction, n=200,k=1(min), and this is about the standard normal distribution.
```{r}
#Use the newly modified functions to plot the probability and density functions for the sample min (N=200)
# CDF
curve(
  porder_anyOrderDistri(x,1,"norm",mean=0,sd=1),
  ylab = "probability",
  xlim = c(-6,1),
  main = "CDF Of Standard Normal Distribution's Min"
)
# PDF
curve(
  dorder_anyOrderDistri(x,1,"norm", mean=0,sd=1),
  ylab = "density",
  xlim = c(-6,1),
  main ="PDF Of Standard Normal Distribution's Min"
)
```
\
The first graphic is the CDF of standard normal distribution' min and the second graphic is the PDF of standard normal distribution' min
