---
title: "Title of Blog Post Deliverable"
author: 'Xinyu Gao'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
    toc: yes
    number_sections: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction 

This assignment is going to explore how to use analytic and simulation ways to develop functions for the distribution of order statistics. 

# Methods
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(tgsify))
```

## Question 1
$$f_k(x)=k{n \choose k}[F(x)]^{k−1}[1−F(x)]^{n−k}f(x)$$
Based on the question introduction, n=200,k=100(median), and this is about the standard normal distribution
```{r}
# according to the above formula (k = 100, n = 200)
dorder <- function(x){
  100*
    choose(200,100)*
    (pnorm(x, mean=0, sd =1))^(100-1)*
    (1-pnorm(x, mean=0, sd =1))^(200-100)*
    dnorm(x, mean=0, sd =1)
}
# plot of the function
curve(
  dorder(x),
  -1,
  1,
  ylab = "Density",
  main = "PDF Of Standard Normal Distribution's Median"
)
```

## Question 2
$$F_{k}(x)=\sum_{j=k}^{N}{n \choose j}[F(x)]^j[1−F(x)]^{n−j}$$
Based on the question introduction, n=200,k=100, and this is about the standard normal distribution
```{r}
# according to the above formula(k = 100, n = 200)
porder <- function(x){
  pbinom(100-1, 200, pnorm(x, mean=0, sd=1), lower.tail = FALSE)
}
curve(
  porder(x),
  -1,
  1,
  ylab = "Probability",
  main = "CDF Of Standard Normal Distribution's Median"
)
```

## Question3
"qorder" is the inverse of "porder" 
\
Step1: create a look up table with value and its corresponding probability. The probability is calculated using "porder"
\
Step2: find the approximate function of the relationship between probability and value using "approxfun". In this function, probability is the input and value is the output
```{r}
qorder <- function(x){
  # look up table
  lt <- data.frame(x=seq(-4,15,by=0.01)) %>%
  mutate(p = porder(x)) 
  # function of the relationship between probability and value
  qf <-approxfun(lt$p,lt$x,ties="ordered")
  # using this function to find corresponding value given probability x
  qf(x)
}

curve(qorder(x),
      0,
      1,
      ylab= "x",
      xlab="quantile",
      main = "Quantile Of Standard Normal Distribution's Median")
```

## Question 4
Use "ecdf" and "porder" function to finish this question
```{r}
set.seed(1)
bmi <- rnorm(200, mean=0, sd=1) %>% median 
# the output is a function
bmi_ecdf <- ecdf(bmi)  
plot(bmi_ecdf, do.points = FALSE, col = "red", lwd = 3, main="ecdf VS CDF",xlim=c(-0.3,0.3))
# use porder to get the plot of the CDF
curve(porder(x),lwd = 3, col = "blue", add=TRUE)

# define the property of the legend
legend(
    "topleft"
  , c("ECDF","CDF")
  , lwd = 3
  , col = c("red","blue")
  , bty = "n"
)
```

## Question 5
Generate histogram of median sampling distribution and use "dorder" to get the theoretical PDF
```{r}
# get 10000 sample
median_norm <- rep(NA,10000)
for (i in seq_along(median_norm))
{
  # calculate median for each sample
  median_norm[i] = rnorm(200,mean = 0, sd=1) %>% median
}
# plot 10000 sample
hist(
  median_norm,
  breaks = 100,
  main = "Sampling distribution VS PDF",
  xlab = "x",
  freq = FALSE
)
# use "dorder" to get the plot of PDF
curve(
  dorder(x),
  lwd = 3, 
  col = "blue",
  add=TRUE
)
```

## Question 6
```{r}
# get 10000 sample
median_norm_2 <- rep(NA,10000)
for (i in seq_along(median_norm_2)){
  # calculate median for each sample
  median_norm_2[i]  <- rnorm(200,mean=0, sd=1) %>% median
}
# sort the list
median_norm_2 <- median_norm_2 %>% sort()
# want to explore 199 quantiles except 0 and 1 quantile
y <- rep(NA, 199)
for (i in seq_along(y))
{
  y[i] = median_norm_2[10000*i/200]
}
# compatible with x
x = qorder((1:199)/200)

plot(x,y
     ,asp=1
     ,main = "Median Theoretical Quantile VS Sample Quantile"
     ,xlab = "Theoretical quantile"
     ,ylab = "Sample quantile"
     )
abline(0,1)

```
\
Based on the above graphic, the plotted points almost fall along the line y=x. Thus the simulated median agree with the theoretical sampling distribution

## Question 7
The updated function is the version of corresponding function from Question 1 using k to replace 100.
```{r}
# n=200,k=k
dorder_anyOrder <- function(x,k){
  k*
    choose(200,k)*
    (pnorm(x, mean=0, sd =1))^(k-1)*
    (1-pnorm(x, mean=0, sd =1))^(200-k)*
    dnorm(x, mean=0, sd =1)
}

# n=200,k=k
porder_anyOrder <- function(x,k){
  pbinom(k-1, 200, pnorm(x, mean=0, sd=1), lower.tail = FALSE)
}

# n = 200,k=k
qorder_anyOrder <- function(x,k){
  lt <- data.frame(x=seq(-4,15,by=0.01)) %>%
     mutate(p = porder_anyOrder(x,k)) 
  qf <-approxfun(lt$p,lt$x, ties ="ordered")
  qf(x)
}
```

## Question 8
```{r}
# get 10000 samples
max_norm <- rep(NA,10000)
for (i in seq_along(max_norm)){
  # calculate the max for each sample
  max_norm[i] = rnorm(200,mean=0, sd=1) %>% max
}
# sort the list
max_norm <- max_norm %>% sort()

# want to explore 199 quantiles
y <- rep(NA,199)
for (i in seq_along(y)){
  y[i] = max_norm[10000*i/200]
}
# use function "qorder_anyOrder" with k = 200(max)
x<-qorder_anyOrder((1:199)/200,200)
plot(x,y, asp = 1,
     xlab = "Theoretical quantile", 
     ylab = "Sample quantile",
     main =  "Max Theoretical Quantile VS Sample Quantile")
abline(0,1)
```
\
Based on the above graphic, the plotted points almost fall along the line y=x. Thus the simulated max agree with the theoretical sampling distribution.

## Question 9
The new function needs two more variables "dist" and  "...". "dist" contains the name information of the distribution and "..." contains the parameter information of the distribution
```{r}
# combine two strings
`%|%` <- function(a,b) paste0(a,b)

dorder_anyOrderDistri <- function(x,k,dist,...){
  # CDF
  F <- eval(parse(text ="p" %|% dist))
  # PDF
  f <- eval(parse(text ="d" %|% dist))
  k*
    choose(200,k)*
    (F(x, ...))^(k-1)*
    (1-F(x, ...))^(200-k)*
    f(x, ...)
}

porder_anyOrderDistri <- function(x,k,dist,...){
  # CDF
  F <- eval(parse(text ="p" %|% dist))
  pbinom(k-1, 200, F(x, ...), lower.tail = FALSE)
}

qorder_anyOrderDistri <- function(x,k,dist,...){
  lt <- data.frame(x=seq(-4,15,by=0.01)) %>%
     mutate(p = porder_anyOrderDistri(x,k,dist,...)) 
  qf <-approxfun(lt$p,lt$x)
  qf(x)
}
```

## Question 10
Based on the question introduction, n=200,k=1(min), and this is about the standard normal distribution.
```{r}
#Use the newly modified functions to plot the probability and density functions for the sample min (N=200)
# CDF
curve(
  porder_anyOrderDistri(x,1,"norm",mean=0,sd=1),
  ylab = "probability",
  xlim = c(-6,1),
  main = "CDF Of Standard Normal Distribution's Min"
)
# PDF
curve(
  dorder_anyOrderDistri(x,1,"norm", mean=0,sd=1),
  ylab = "density",
  xlim = c(-6,1),
  main ="PDF Of Standard Normal Distribution's Median"
)
```

