---
title: "Title of Blog Post Deliverable"
author: 'Xinyu Gao'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
    toc: yes
    number_sections: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction 

The purpose of this assignment is to study the coverage probability. 

## Background

In this assignment, we will use following knowledge. 1. Given a sample, use MLE to estimate the parameters of this sample's distribution (normal distribution) 2. Generate new samples based on the parameters calculated from the previous step and find the median of these samples 3. Plot all the median and calculate the confidence interval of the median distribution 4. Compare the results from step3 with the median of the original sample and then calculate the coverage probability. 

# Methods

```{r}
suppressPackageStartupMessages(require(tidyverse))
suppressPackageStartupMessages(require(dplyr))
suppressPackageStartupMessages(require(stats4))
```

## Step1
MLE is a method that tries a lot of possible parameters and find the parameters that can maximum the probability of obtaining the sample data. In other word, we need to find parameters that can maximum the following formula $$ f(x_1,x_2,...,x_n|parameters) $$ In R, we need to use mle to achieve this task. mle needs two primary parameters:1. negative likelihood function which needs to be minimized 2. starting point for the coefficient vector.
```{r}
# generate a single sample
set.seed(1)
sample <- rnorm(201,mean=0,sd=1)
# use mle to calculate the parameters based on the sample
ll_norm <- function(mean,sd){
  -sum(dnorm(sample,mean = mean,sd = sd,log=T))
}
z_norm <- mle(minuslogl=ll_norm, start=list(mean=1,sd=1),
                 method = "L-BFGS-B",
                 lower = c(0,0.01))
mean <- coef(z_norm)[1]
sd <- coef(z_norm)[2]
print(paste("The estimated mean of normal distribution is: ",mean))
print(paste("The estimated standard deviation of normal distribution is: ", sd))
```

## Step2
```{r}
# I want to get 10000 samples and for each sample, there are 100 values
result <- rep(NA, 10000)
for (i in seq_along(result))
{
  # find the median of each sample using quantile function
  result[i] = quantile(rnorm(100,mean = mean, sd = sd),0.5)
}
hist(result,freq=FALSE,xlab = "Median",main="Distribution of Median")
```
\
This is the histogram of the sampling distribution of the median. The shape of the distribution is very close to the normal distribution.

## Step3
Based on the graphic from Step2, the sampling distribution of median is very close to normal distribution, which is symmetric. Thus, if we want to get the 95% confidence interval, we can use symmetric density interval method, which is to remove the leftmost 0.025% and rightmost 0.025% values in the distribution.
```{r}
# 95%
x_0.025 = quantile(result,0.025)
x_0.975 = quantile(result,0.975)
hist(result,freq=FALSE,xlab = "Median",main="Distribution of Median")
abline(v=x_0.025,col="red")
abline(v=x_0.975,col="red")
print(paste("The minimum value among 95% confidence interval is: ",x_0.025))
print(paste("The maximum value among 95% confidence interval is: ",x_0.975))
```
\
This is the histogram of the sampling distribution of the median. The red lines represent the range of the 95% confidence interval and the blue line represents the ture median of the original sample.

## Step4 and Step5
The coverage probability is calculated by comparing the estimated population parameters from samples (using confidence interval) with the true parameters of the population. Among all the samples, if the estimated population parameters from samples capture the true parameters of the population, the sample should be recorded. The coverage probability is the proportion of samples for which the confidence interval captured the true population parameter.
```{r}
# function that can return the 95% confidence interval of the sampling distribution of the median
sample_95_confidence <- function()
{
  sample <- rnorm(201,mean=0,sd=1)
  # use mle to calculate the parameters based on the sample
  ll_norm <- function(mean,sd){
    -sum(dnorm(sample,mean = mean,sd = sd,log=T))
  }
  z_norm <- mle(minuslogl=ll_norm, start=list(mean=1,sd=1),
                 method = "L-BFGS-B",
                 lower = c(0,0.01))
  mean <- coef(z_norm)[1]
  sd <- coef(z_norm)[2]
  # 5000 samples
  result <- rep(NA, 5000)
  for (i in seq_along(result))
  {
    # each sample has 500 values and find the median of each sample using quantile
    result[i] = quantile(rnorm(500,mean = mean, sd = sd),0.5)
  }
  # find the minimum value among the 95% confidence interval
  x_0.025 = quantile(result,0.025)
  # fine the maximum value among the 95% confidence interval
  x_0.975 = quantile(result,0.975)
  return(c(x_0.025, x_0.975))
}

# get 1000 sampling distributions
num_sample_dis <- 1000
result <- replicate(num_sample_dis,sample_95_confidence())
result <- as.data.frame(t(result))
result$index <- 1:dim(result)[1]
colnames(result) <- c("low","up","index")
low <- result %>%
  select(low,index) %>%
  rename(value = low)
up <- result %>%
  select(up,index) %>%
  rename(value = up)
low_up <- rbind(low,up)
low_up['Status'] = 0
# calculate the coverage probability
count <- 0 

for (i in 1:num_sample_dis)
{
  # get the minimum value and maximum value among the 95% confidence interval
  values <- low_up %>%
    filter(index == i) %>%
    pull(value)
  # the sample for which the confidence interval captured the population parameter
  if (0 >= values[1] & 0 <= values[2])
  {
    count <- count +1
    low_up['Status'][low_up['index']==i] = 1
  }
  cov_prob <- count/num_sample_dis
}
print(paste("The coverage probability is: ",cov_prob))

# get maximum 5 sampling distributions that cover the population parameters
cover <- low_up %>%
  filter(Status == 1) %>%
  group_by(index) %>%
  summarise(Value = value,Status = Status) %>%
  ungroup() %>%
  slice(1:10)

# get maximum 5 sampling distributions that don't cover the population parameters
not_cover <- low_up %>%
  filter(Status == 0) %>%
  group_by(index) %>%
  summarise(Value = value,Status = Status) %>%
  ungroup() %>%
  slice(1:10)

# combine them and modify the index to make the furure more clear
combine = rbind(cover,not_cover)
total_len = dim(unique(combine['index']))[1]
index <- unique(combine['index']) %>% `[[`(1)
for (i in 1:total_len)
{
  combine['index'][combine['index'] == index[i]] = -i
}

#plot the graphic
combine$Status <- as.factor(combine$Status)
combine %>%
  ggplot(aes(x=Value,y=-index,group = index, color = Status))+
  geom_line()+
  #scale_color_manual(values = c("blue","red"))+
  geom_vline(xintercept = 0,col="grey")+
  labs(
    x = "Pop.parameter of interest",
    y = "Sample"
  )+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(),legend.position = "none")
```
\
The coverage probability is calculated by comparing the estimated population parameters from samples (using confidence interval) with the true parameters of the population.The coverage probability is the proportion of samples for which the confidence interval captured the true population parameter. In this case, the coverage probability is 0.939, which means among 1000 sampling distribution, there are about 60 distributions that don't capture the true population parameter like the red lines in the graphic.

## Step6
We can increase the original sample size to see whether the estimated parameters using the MLE will be more close to the real parameters. Also, we can try other methods, like MM to estimate the parameters. When generating the sampling distribution of median, I tried 10000 samples and for each sample, there was 100 values. We can adjust these two numbers to see if there is any difference.

