---
title: "Title of Blog Post Deliverable"
author: 'Xinyu Gao'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
    toc: yes
    number_sections: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction 

The purpose of this assignment is to answer a series of questions related to the transformation of data using analytic or simulation method. 

## Background

To finish this assignment, we need to rely on some functions. "pgamma","plnorm", and "punif" are required to generate the CDF figure; "dgamma", "dlnorm", and "dunif" are required to generate the PDF figure; "rgamma", "rlnorm", and "runif" are required to generate random variables; "qgamma","qlnorm", and "qunif" are required to get value based on a given quantile.

# Methods

```{r}
suppressPackageStartupMessages(require(tidyverse))
suppressPackageStartupMessages(require(ggplot2))
```

## Part1

### Question 1

#### Gamma
Gamma mean: $$shape*scale$$
```{r}
x <- seq(-1,10,by=0.01)
# PDF
pdf <- dgamma(x,shape = 3, scale = 1)
# CDF 
cdf <- pgamma(x,shape = 3, scale = 1)

mean <- 3*1
median <- qgamma(0.5, shape = 3, scale=1)

#PDF
plot(x,pdf, main="Gamma Distribution's PDF", type ="l")
abline(v = mean, col = "red")
abline(v = median, col = "blue")
legend("right", c("mean","median"),col =  c("red","blue"), lty=c(1,1))

#CDF
plot(x,cdf, main="Gamma Distribution's CDF", type ="l")
abline(v = mean, col = "red")
abline(v = median, col = "blue")
legend("right", c("mean","median"),col =  c("red","blue"), lty=c(1,1))
```

#### logNormal
logNormal mean: $$	{\displaystyle \exp \left(\mu +{\frac {\sigma ^{2}}{2}}\right)}$$
```{r}
x <- seq(-1,10,by=0.01)
# PDF
pdf <- dlnorm(x,-1,1)
# CDF
cdf <- plnorm(x,-1,1)
mean <- exp(-1+1^2/2)
median <- qlnorm(0.5, -1,1)

#PDF
plot(x,pdf, main="LogNormal Distribution's PDF", type ="l")
abline(v = mean, col = "red")
abline(v = median, col = "blue")
legend("right", c("mean","median"),col =  c("red","blue"), lty=c(1,1))

#CDF
plot(x,cdf, main="LogNormal Distribution's CDF", type ="l")
abline(v = mean, col = "red")
abline(v = median, col = "blue")
legend("right", c("mean","median"),col =  c("red","blue"), lty=c(1,1))
```

#### Uniform
Uniform mean:$${\tfrac {1}{2}}(min+max)$$
```{r}
x <- seq(-1,20,by=0.01)
# PDF
pdf <- dunif(x,0,12)
# CDF
cdf <- punif(x,0,12)
mean <- (0+12)/2
median <- qunif(0.5, 0,12)

#PDF
plot(x,pdf, main="Uniform Distribution's PDF", type ="l")
abline(v = mean, col = "red")
abline(v = median, col = "blue")
legend("right", c("mean","median"),col =  c("red","blue"), lty=c(1,1))

#CDF
plot(x,cdf, main="Uniform Distribution's CDF", type ="l")
abline(v = mean, col = "red")
abline(v = median, col = "blue")
legend("right", c("mean","median"),col =  c("red","blue"), lty=c(1,1))
```
\
For the uniform distribution, the mean and the median are the same. Thus the two lines overlap.

### Question 2

#### Gamma
```{r}
x <- rgamma(10000, shape = 3, scale =1)

# PDF log transformation
Y <- log(x)
plot(density(Y),type = "l",main ="Gamma Distribution's log transformation PDF", xlab = "x")
abline(v = mean(Y), col = "red")
abline(v = median(Y), col = "blue")
legend("left",c("mean","median"), col = c("red","blue"), lty=c(1,1))

# CDF log transformation
plot(ecdf(Y), main = "Gamma Distribution's log transformation CDF")
abline(v = mean(Y), col = "red")
abline(v = median(Y), col = "blue")
legend("left",c("mean","median"), col = c("red","blue"), lty=c(1,1))
```

#### LogNormal
```{r}
x <- rlnorm(10000, -1, 1)

# PDF log transformation
Y <- log(x)
plot(density(Y),type = "l",main ="LogNormal Distribution's log transformation PDF",xlab = "x")
abline(v = mean(Y), col = "red")
abline(v = median(Y), col = "blue")
legend("left",c("mean","median"), col = c("red","blue"), lty=c(1,1))

# CDF log transformation
plot(ecdf(Y), main = "LogNormal Distribution's log transformation CDF")
abline(v = mean(Y), col = "red")
abline(v = median(Y), col = "blue")
legend("left",c("mean","median"), col = c("red","blue"), lty=c(1,1))
```
\
For the logNormal distribution, the mean and the median are the same. Thus the two lines overlap.

#### Uniform
```{r}
x <- runif(10000, 0, 12)

# PDF log transformation
Y <- log(x)
plot(density(Y),type = "l",main ="Uniform Distribution's log transformation PDF",xlab = "x")
abline(v = mean(Y), col = "red")
abline(v = median(Y), col = "blue")
legend("left",c("mean","median"), col = c("red","blue"), lty=c(1,1))

# CDF log transformation
plot(ecdf(Y), main = "Uniform Distribution's log transformation CDF")
abline(v = mean(Y), col = "red")
abline(v = median(Y), col = "blue")
legend("left",c("mean","median"), col = c("red","blue"), lty=c(1,1))
```

### Question 3
Arithmetic mean $$\frac{\left(x_{1}+x_{2}+...+x_{n}\right)}{n}$$
Geometric mean $$\left(x_{1}\cdot x_{2}\cdot...\cdot x_{n}\right)^{\frac{1}{n}}$$
How to calculate geometric mean $$e^{\frac{\left(\log x_{1}+\log x_{2}+...+\log x_{n}\right)}{n}}=e^{\left(\frac{1}{n}\cdot\log\left(x_{1}x_{2}...x_{n}\right)\right)}=e^{\log\left(\left(x_{1}\cdot x_{2}\cdot...\cdot x_{n}\right)^{\frac{1}{n}}\right)}=\left(x_{1}\cdot x_{2}\cdot...\cdot x_{n}\right)^{\frac{1}{n}}$$

#### Gamma
```{r}
set.seed(2)
# arithmetic mean
gamma_arthi <- data.frame(arthmetic = replicate(1000,mean(rgamma(100,shape=3,scale=1))))
# geometric mean
gamma_geo <- data.frame(geometric = replicate(1000,exp(mean(log(rgamma(100,shape=3,scale=1))))))
# generate the graphic
cbind(gamma_arthi,gamma_geo) %>%
  ggplot(aes(x=arthmetic,y=geometric)) +
  geom_point()+
  geom_smooth(method = "lm")+
  labs(
    x= "Arthmetic Mean",
    y= "Geomeric Mean",
    title ="Gamma Distribution Arthmetic Mean VS Geomeric Mean "
  )
```

#### LogNormal
```{r}
set.seed(2)
# arithmetic mean
logNormal_arthi <- data.frame(arthmetic = replicate(1000,mean(rlnorm(100,-1,1))))
# geometric mean
logNormal_geo <- data.frame(geometric = replicate(1000,exp(mean(log(rlnorm(100,-1,1))))))
# generate the graphic
cbind(logNormal_arthi,logNormal_geo) %>%
  ggplot(aes(x=arthmetic,y=geometric)) +
  geom_point()+
  geom_smooth(method = "lm")+
  labs(
    x= "Arthmetic Mean",
    y= "Geomeric Mean",
    title ="LogNormal Distribution Arthmetic Mean VS Geomeric Mean "
  )
```

#### Uniform
```{r}
set.seed(2)
# arithmetic mean
uniform_arthi <- data.frame(arthmetic = replicate(1000,mean(runif(100,0,12))))
# geometric mean
uniform_geo <- data.frame(geometric = replicate(1000,exp(mean(log(runif(100,0,12))))))
# generate the graphic
cbind(uniform_arthi,uniform_geo) %>%
  ggplot(aes(x=arthmetic,y=geometric)) +
  geom_point()+
  geom_smooth(method = "lm")+
  labs(
    x= "Arthmetic Mean",
    y= "Geomeric Mean",
    title ="Uniform Distribution Arthmetic Mean VS Geomeric Mean "
  )
```


### Question 4
```{r}
# set seed
set.seed(2)
# gamma arithmetic mean
gamma_arthi_value <- replicate(1000,mean(rgamma(100,shape=3,scale=1)))
# gamma geometric mean
gamma_geo_value <- replicate(1000,exp(mean(log(rgamma(100,shape=3,scale=1)))))
# find difference
gamma_diff = gamma_arthi_value-gamma_geo_value
# draw the figure
data.frame(gamma_diff) %>%
  ggplot(aes(x=gamma_diff))+
  geom_histogram(bins=30,
                 col = "black",
                 alpha=0.5,
                 breaks = seq(0,1.2,by = 0.05)) +
  labs(
    x="Gamma Difference",
    y="Count"
  )

# logNormal arithmetic mean
logNormal_arthi_value <- replicate(1000,mean(rlnorm(100,-1,1)))
# logNormal geometric mean
logNormal_geo_value <- replicate(1000,exp(mean(log(rlnorm(100,-1,1)))))
# find the difference 
logNormal_diff = logNormal_arthi_value-logNormal_geo_value
# draw the figure
data.frame(logNormal_diff) %>%
  ggplot(aes(x=logNormal_diff))+
  geom_histogram(bins=30,
                 col = "black",
                 alpha=0.5,
                 breaks=seq(0,0.6,by=0.05))+
  labs(
    x="LogNormal Difference",
    y="Count"
  )

# uniform arithmetic mean
uniform_arthi_value <- replicate(1000,mean(runif(300,0,12)))
# uniform geometric mean
uniform_geo_value <- replicate(1000,exp(mean(log(runif(300,0,12)))))
# find the difference 
uniform_diff = uniform_arthi_value-uniform_geo_value
# draw the figure
data.frame(uniform_diff) %>%
  ggplot(aes(x=uniform_diff))+
  geom_histogram(bins=30,
                 col = "black",
                 alpha=0.5,
                 breaks=seq(0,2.6,by=0.05))+
  labs(
    x="Uniform Difference",
    y="Count"
  )
```
\
For all three figures, the X-axis is the difference between arithmetic mean and geometric mean, the Y-axis id the count of each difference. From all three figures, arithmetic mean is always larger or equal to geometric mean. 
\
Proof of this conclusion: suppose calculate the arithmetic mean and geometric mean of two values a and b. 
\
Arithmetic mean $$\frac{\left(a+b\right)}{2}$$
Geometric mean $$\sqrt{ab}$$
Arithmetic mean^2 - Geometric mean^2
$$\left(\frac{\left(a+b\right)}{2}\right)^{2}\ \ -\ \left(\sqrt{ab}\right)^{2}=\frac{\left(a^{2}+2ab+b^{2}\right)}{4}-\frac{4ab}{4}=\frac{\left(a^{2}-2ab+b^{2}\right)}{4}=\frac{\left(a-b\right)^{2}}{4}\ge0$$

## Part 3
Start from Part 1 Question 1 and 2:
Gamma distribution: $$log(E[x])  = log(3) = 1.09861228867 > E[log(x)] = 0.922784335098 $$
LogNormal distribution: $$log(E[x])  = log(0.606462351047) = −0.500112628292
 > E[log(x)] = -1.00000533767 $$
 Uniform distribution: $$log(E[x])  = log(6) = 1.79175946923
 > E[log(x)] = 1.48490664979$$
Assume Relationship:
$$E[log(X)] ≤  log(E(X))$$
Proof of assumption
\
According to the Jensens inequality: for any concave function f, $$E(f(X)) ≤ f(E(X))$$
A function is concave if it's second derivative is negative
\
Proof f: x -> log(x) is a concave function
$$ f'' = (log(x))'' = (\frac{1}{x})' = -\frac{1}{x^2} <0 $$


